---
title: "DATA 608: Knowledge And Visual Analytics"
author: "Walt Wells, Fall 2017"
subtitle: "HW3 - CDC Mortality: EXPLORATION"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
    self_contained: no
---


# Environment Prep

```{r, message=F, warning=F}
if (!require('dplyr')) install.packages('dplyr')
if (!require('plotly')) install.packages('plotly')
if (!require('googleVis')) install.packages('googleVis')
```

# Data Source

The Underlying Cause of Death data available on WONDER are county-level national mortality and population data spanning the years 1999-2015. Data are based on death certificates for U.S. residents. Each death certificate identifies a single underlying cause of death and demographic data. The number of deaths, crude death rates or age-adjusted death rates, and 95% confidence intervals and standard errors for death rates can be obtained by place of residence (total U.S., region, state and county), age group (single-year-of age, 5-year age groups, 10-year age groups and infant age groups), race, Hispanic ethnicity, gender, year, cause-of-death (4-digit ICD-10 code or group of codes), injury intent and injury mechanism, drug/alcohol induced causes and urbanization categories. Data are also available for place of death, month and week day of death, and whether an autopsy was performed. 

https://wonder.cdc.gov/wonder/help/ucd.html#

# Load Data

```{r}
data <- read.csv('https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture3/data/cleaned-cdc-mortality-1999-2010-2.csv')
```

## Quick exploration

```{r}
dim(data)
#table(data$ICD.Chapter)
#summary(data)
#str(data)
```


# Question 1:
As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.

```{r}
mypick <- "Neoplasms"
sub <- data %>%
    filter(Year == 2010, ICD.Chapter==mypick) %>%
    select(State, Crude.Rate) %>%
    arrange(Crude.Rate)
```

```{r results='asis', tidy=FALSE}
#plot_ly(sub, y=~Crude.Rate, x=~State)
t <- paste0("Cause: ", mypick)
p <- gvisColumnChart(sub, 
                     options=list(
                         title=t, 
                         legend="none"))
print(p, 'chart')
```

# Question 2:
Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population.

```{r}
mypick <- "Certain infectious and parasitic diseases"
mypick <- "Neoplasms"
sub2 <- data %>%
    filter(ICD.Chapter==mypick) %>%
    select(State, Year, Deaths, Population, Crude.Rate) %>%
    group_by(Year) %>%
    mutate(Nat.Avg = round((sum(Deaths) / sum(Population)) * 10^5, 1)) %>%
    select(State, Year, Crude.Rate, Nat.Avg) %>%
    ungroup()

sub2$State <- as.character(sub2$State)
```

```{r}
#state <- "IL"
state <- "UT"
sub3 <- sub2 %>%
    filter(State==state) %>%
    select(Year, Crude.Rate, Nat.Avg) %>%
    rename(State.Avg = Crude.Rate)
```

```{r results='asis', tidy=FALSE}

t <- paste0("Cause: ", mypick,  "  |  State: ", state)
q <- gvisAreaChart(sub3, 
                     options=list(
                         title=t, 
                         hAxis="{format:'####'}"))
print(q, 'chart')
```

```{r results='asis', tidy=FALSE}
Motion=gvisMotionChart(sub2, 
                       idvar="State", 
                       timevar="Year")
print(Motion, 'chart')
```

